Capturing Subdiffusive Solute Dynamics and Predicting Selectivity in Nanoscale Pores with Time Series Modeling

### 2020-05-02 Ben_Manuscripts/stochastic_transport/Final_Draft_JCTC.tex
More comments on draft.

  Highly selective separations membranes are desirable in numerous
  applications. The ability to efficiently separate ions from saline water
  sources using membranes has been actively pursued for years in an effort to
  create potable water for people in water-scarce
  regions.~\cite{werber_materials_2016} Even in relatively safe municipal water
  supplies, there is a need for membranes that can specifically separate
  potentially harmful organic micropollutants such as fertilizers, pesticides,
  pharmaceuticals and personal care products.~\cite{barbosa_occurrence_2016} 
  %BJCB: could cut this last sentence. It's kind of awkward  
  In the materials industry, there is a strong interest in creating breathable
  fabrics which selectively allow passage of water
  vapor.~\cite{mondloch_destruction_2015}

@@

  % BJCB: split into two paragraphs  
  In this work, we study the inverted hexagonal, H\textsubscript{II}, LLC phase. 
  The H\textsubscript{II} phase has densely packed, uniform-sized pores on the 
  order of 1 nm in size. A perfectly aligned H\textsubscript{II} phase has 
  an ideal geometry for high throughput separations.~\cite{feng_scalable_2014,feng_thin_2016}
  %BJCB: could add a sentence saying that in pracftice, alignment is hard. Not sure if worth acknowledging for this paper.

@@

  %MRS: some of these discussions might belong in the SI, with just the top line information in the main text.  People don't really care about the details of the individual parameters, just what story that tells.
  %BJCB: I think the figures need to stay though   
  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{1mode_AD_hops.pdf}
  \caption{}\label{fig:1mode_AD_hops}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{1mode_AD_dwells.pdf}
  \caption{}\label{fig:1mode_AD_dwells}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{hurst_barchart.pdf}
  \caption{}\label{fig:hurst_barchart}
  \end{subfigure}
  \caption{The parameters of the one mode model reveal differences in dynamics
	  between solutes. (a) We parameterized Gaussian,
	  $\mathcal{N}(\sigma)$, and L\'evy stable, $L(\sigma, \alpha_h)$,
	  distributions to describe solute hop lengths. We assume the mean
	  ($\mu$) to be zero for these distributions and no
	  skewness ($\beta = 0$) in the L\'evy stable distributions. High
	  values of $\sigma$ and lower values of $\alpha_h$ result in larger
	  hops. (b) We parameterized a pure power law, $P(\alpha)$, and a
	  truncated power law, $P_T(\alpha, \lambda)$, distribution to describe
	  solute dwell times. Lower values of $\alpha$ lead to heavier power
	  law tails and higher values of $\lambda$ truncate the distribution at
	  lower dwell times. (c) Finally, we parameterized the hop
	  autocorrelation function, $\gamma(H)$, to describe the degree of
	  correlation between hops. Simulations with higher values of $H$ display
	  behavior closer to the Brownian limit.}\label{fig:1mode_parameters}
  \end{figure}

@@

  % BJCB: I might be able to get away with moving this paragraph to SI if I briefly
  % mention the reason in the above paragraph.
  The two mode models display curvature more consistent with MD but for
  non-physical reasons. Every time a switch between the pores and tails occurs,
  the width of the distribution used to model hops changes. We are unaware of
  an appropriate technique which can correlate hops that come from different
  hop distributions, therefore every time a mode switch occurs, the correlation
  structure is broken. Solutes that switch between modes the least show
  predicted MSDs with the greatest curvature. Due to the much larger
  accessible volume that a smaller molecule has, methanol spends $>90\%$ of its
  time in the hydrophobic tails (see Figure~\ref{fig:AD_mode_occupation}), so
  mode transitions are relatively rare and the predicted MSDs have significant
  curvature. This artificially solves the problem of long timescale
  correlation, however we do not recommend this as it has no physical basis.

@@

  %BJCB: maybe move the next two paragraphs to the SI?
  Examining first the parameters of the one mode model, we can begin to break
  down the trends in solute MSDs. The parameters belonging to ethylene glycol
  and methanol are relatively similar, which is consistent with their similar MD
  MSDs and qualitatively similar MD time series.
  Relative to ethylene glycol, methanol tends to stay trapped for less
  time and takes larger hops but the most substantial difference is with
  respect to their Hurst parameters. Methanol has the lowest $H$ of all the
  solutes studied because it spends the majority of its time outside the pore
  region where collisions with tails are frequent.
  Urea has the third highest MSD which is primarily a consequence of more
  frequent and longer dwell times (lower $\alpha$ and $\lambda$). Urea's hop
  lengths ($\sigma$) and correlation ($H$) are comparable to ethylene glycol
  and methanol. Acetic acid has the smallest MSD among the solutes studied due
  to longer periods of entrapment and shorter hops. Its trapping behavior is
  parameterized by an $\alpha$ value significantly lower than other solutes,
  but an intermediate $\lambda$ value, suggesting it experiences many
  medium-length periods of entrapment. Its hops are smaller but are slightly
  compensated by a heavier tailed distribution (lower $\alpha_h$) than the
  other solutes.

@@

  Urea spends the largest fraction of its time trapped via association with
  sodium ions. It does so 31\% of the total time while in the tails (state 3)
  and 21\% of the total time while in the pores (state 7). Note that sodium
  does not drift significantly into the tails but sits close to the pore/tail
  region boundary. The electron-dense and unshielded oxygen atom of urea's
  carbonyl group is prone to associate with positively charged sodium ions. The
  nitrogen atoms of urea are only weak hydrogen bond donors. Therefore, the
  transport rate of urea is likely to be most significantly modified by
  removing or changing the identity of the counter-ion.

  Ethylene glycol spends the largest fraction of its time trapped in a hydrogen
  bonded state. It does so 24\% of the total time while in the tails (state 2)
  and 32\% of the total time while in the pores (state 6). The two hydroxyl
  groups of ethylene glycol readily donate their hydrogen atoms to the
  carboxylate head groups and the ether linkages between the head groups and
  monomer tails. The transport rate of ethylene glycol might be modified by
  removing hydrogen bond accepting groups from the LLC monomers, especially
  those which stabilize the solute in the tails (i.e. the ether linkages).

  Methanol spends most of its time unbound in the tail region (state 1) and
  spends a significant portion of time hydrogen bonded while in the tail
  region. Tail region hydrogen bonds are donated from methanol to the ether
  linkages between the monomer head groups and tails. One might modify the
  rate of methanol transport by controlling its partition into the monomer
  tails. This might be achieved by adding cross-linkable groups near the
  monomer head groups.

  Finally, acetic acid spends the majority of its time hydrogen bonding both in
  and out of the pore (states 2 and 6). Although it has an unshielded carbonyl
  group in its structure, association with sodium ions in this environment is
  apparently a much weaker interaction than hydrogen bonding.  As with ethylene
  glycol, one might modify the transport rate of acetic acid by removing
  hydrogen bond accepting constituents of the LLC monomer. With this
  modification, we hypothesize that acetic acid might show similar transport
  rates to urea given their structural similarity.

@@

  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{H_v_state.pdf}
  \caption{}\label{fig:H_v_state}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{sigma_v_state.pdf}
  \caption{}\label{fig:sigma_v_state}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{alpha_v_state.pdf}
  \caption{}\label{fig:alpha_v_state}
  \end{subfigure}
  \caption{The parameters of the MSDDM are strong functions of trapping
	  mechanisms. We observe different parameters but with similar trends
	  between the tail and pore region. The states are defined in
	  Table~\ref{table:states}. See Figure~\ref{fig:state_probabilities}
	  for a description of the abbreviation under each state number. The
	  legend in (a) applies to all subplots. (a) Motion is highly
	  anti-correlated in trapped states. As the number of simultaneously
	  influencing trapping mechanisms increases, the Hurst parameter, $H$,
	  decreases. $H$ is highest (closest to Brownian) during transitions 
	  between states (state T). (b) As more trapping mechanisms simultaneously influence solutes,
	  the width of the hop length distribution ($\sigma$) decreases. The
	  largest hops occur when solutes are unbound in the pores. (c) The
	  weight of the hop length distribution's tails, parameterized by
	  $\alpha_h$, increases as more trapping mechanisms influence solutes
	  simultaneously. The transitional hop distributions have among the
	  heaviest tails.}\label{fig:msddm_parameters}
  \end{figure}

>>>>

  Highly selective separations membranes are desirable in numerous
  applications. The ability to efficiently separate ions from saline water
  sources using membranes has been actively pursued for years in an effort to
  create potable water for people in water-scarce
  regions.~\cite{werber_materials_2016} Even in relatively safe municipal water
  supplies, there is a need for membranes that can specifically separate
  potentially harmful organic micropollutants such as fertilizers, pesticides,
  pharmaceuticals and personal care products.~\cite{barbosa_occurrence_2016} 
  %BJCB: could cut this last sentence. It's kind of awkward  
  %MRSB: agree  
  %In the materials industry, there is a strong interest in creating breathable
  %fabrics which selectively allow passage of water
  %vapor.~\cite{mondloch_destruction_2015}

@@

  % BJCB: split into two paragraphs  
  In this work, we study the inverted hexagonal, H\textsubscript{II}, LLC phase. 
  The H\textsubscript{II} phase has densely packed, uniform-sized pores on the 
  order of 1 nm in size. A perfectly aligned H\textsubscript{II} phase has 
  an ideal geometry for high throughput separations.~\cite{feng_scalable_2014,feng_thin_2016}
  %BJCB: could add a sentence saying that in pracftice, alignment is hard. Not sure if worth acknowledging for this paper.
  %MRSB: nah, not here.

@@

  %MRS: some of these discussions might belong in the SI, with just the top line information in the main text.  People don't really care about the details of the individual parameters, just what story that tells.
  %BJCB: I think the figures need to stay though   
  %MRSB: Hmm. I'm not entirely convinced, but we can leave it in for now and see what the reviewers think.
  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{1mode_AD_hops.pdf}
  \caption{}\label{fig:1mode_AD_hops}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{1mode_AD_dwells.pdf}
  \caption{}\label{fig:1mode_AD_dwells}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{hurst_barchart.pdf}
  \caption{}\label{fig:hurst_barchart}
  \end{subfigure}
  \caption{The parameters of the one mode model reveal differences in dynamics
	  between solutes. (a) We parameterized Gaussian,
	  $\mathcal{N}(\sigma)$, and L\'evy stable, $L(\sigma, \alpha_h)$,
	  distributions to describe solute hop lengths. We assume the mean
	  ($\mu$) to be zero for these distributions and no
	  skewness ($\beta = 0$) in the L\'evy stable distributions. High
	  values of $\sigma$ and lower values of $\alpha_h$ result in larger
	  hops. (b) We parameterized a pure power law, $P(\alpha)$, and a
	  truncated power law, $P_T(\alpha, \lambda)$, distribution to describe
	  solute dwell times. Lower values of $\alpha$ lead to heavier power
	  law tails and higher values of $\lambda$ truncate the distribution at
	  lower dwell times. (c) Finally, we parameterized the hop
	  autocorrelation function, $\gamma(H)$, to describe the degree of
	  correlation between hops. Simulations with higher values of $H$ display
	  behavior closer to the Brownian limit.}\label{fig:1mode_parameters}
  \end{figure}

@@

  % BJCB: I might be able to get away with moving this paragraph to SI if I briefly
  % mention the reason in the above paragraph.
  The two mode models display curvature more consistent with MD but for
  non-physical reasons. Every time a switch between the pores and tails occurs,
  the width of the distribution used to model hops changes, and the correlation structure is broken. 
  %MRSB: I cut below.
  %We are unaware of
  %an appropriate technique which can correlate hops that come from different
  %hop distribution every time a mode switch occurs, the correlation
  %structure is broken. 
  Solutes that switch between modes the least show
  predicted MSDs with the greatest curvature. Due to the much larger
  accessible volume that a smaller molecule has, methanol spends $>90\%$ of its
  time in the hydrophobic tails (see Figure~\ref{fig:AD_mode_occupation}), so
  mode transitions are relatively rare and the predicted MSDs have significant
  curvature. This artificially resolves the problem of long timescale
  correlation, however, it has no physical basis.

@@

  %BJCB: maybe move the next two paragraphs to the SI?
  %MRSB: I think that this is more useful in the main text, since it connects models to the simulations and the physics.
  Examining first the parameters of the one mode model, we can begin to break
  down the trends in solute MSDs. The parameters belonging to ethylene glycol
  and methanol are relatively similar, which is consistent with their similar MD
  MSDs and qualitatively similar MD time series.
  Relative to ethylene glycol, methanol tends to stay trapped for less
  time and takes larger hops but the most substantial difference is with
  respect to their Hurst parameters. Methanol has the lowest $H$ of all the
  solutes studied because it spends the majority of its time outside the pore
  region where collisions with tails are frequent.
  Urea has the third highest MSD which is primarily a consequence of more
  frequent and longer dwell times (lower $\alpha$ and $\lambda$). Urea's hop
  lengths ($\sigma$) and correlation ($H$) are comparable to ethylene glycol
  and methanol. Acetic acid has the smallest MSD among the solutes studied due
  to longer periods of entrapment and shorter hops. Its trapping behavior is
  parameterized by an $\alpha$ value significantly lower than other solutes,
  but an intermediate $\lambda$ value, suggesting it experiences many
  medium-length periods of entrapment. Its hops are smaller but are slightly
  compensated by a heavier tailed distribution (lower $\alpha_h$) than the
  other solutes.

@@

  %MRSB: I think people like the physical intuition gained. I added bullet points to aid reading. 
  \begin{itemize}
  \item \textit{Urea} spends the largest fraction of its time trapped via association with
  sodium ions. It does so 31\% of the total time while in the tails (state 3)
  and 21\% of the total time while in the pores (state 7). Note that sodium
  does not drift significantly into the tails but sits close to the pore/tail
  region boundary. The electron-dense and unshielded oxygen atom of urea's
  carbonyl group is prone to associate with positively charged sodium ions. The
  nitrogen atoms of urea are only weak hydrogen bond donors. Therefore, the
  transport rate of urea is likely to be most significantly modified by
  removing or changing the identity of the counter-ion.

  \item \textit{Ethylene glycol} spends the largest fraction of its time trapped in a hydrogen
  bonded state. It does so 24\% of the total time while in the tails (state 2)
  and 32\% of the total time while in the pores (state 6). The two hydroxyl
  groups of ethylene glycol readily donate their hydrogen atoms to the
  carboxylate head groups and the ether linkages between the head groups and
  monomer tails. The transport rate of ethylene glycol might be modified by
  removing hydrogen bond accepting groups from the LLC monomers, especially
  those which stabilize the solute in the tails (i.e. the ether linkages).

  \item \textit{Methanol} spends most of its time unbound in the tail region (state 1) and
  spends a significant portion of time hydrogen bonded while in the tail
  region. Tail region hydrogen bonds are donated from methanol to the ether
  linkages between the monomer head groups and tails. One might modify the
  rate of methanol transport by controlling its partition into the monomer
  tails. This might be achieved by adding cross-linkable groups near the
  monomer head groups.

  \item \textit{Acetic acid} spends the majority of its time hydrogen bonding both in
  and out of the pore (states 2 and 6). Although it has an unshielded carbonyl
  group in its structure, association with sodium ions in this environment is
  apparently a much weaker interaction than hydrogen bonding.  As with ethylene
  glycol, one might modify the transport rate of acetic acid by removing
  hydrogen bond accepting constituents of the LLC monomer. With this
  modification, we hypothesize that acetic acid might show similar transport
  rates to urea given their structural similarity.
  \end{itemize}

@@

  %MRSB: discussion of parameter specifics is a potential place to move.
  \begin{figure}
  \centering
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{H_v_state.pdf}
  \caption{}\label{fig:H_v_state}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{sigma_v_state.pdf}
  \caption{}\label{fig:sigma_v_state}
  \end{subfigure}
  \begin{subfigure}{0.325\textwidth}
  \includegraphics[width=\textwidth]{alpha_v_state.pdf}
  \caption{}\label{fig:alpha_v_state}
  \end{subfigure}
  \caption{The parameters of the MSDDM are strong functions of trapping
	  mechanisms. We observe different parameters but with similar trends
	  between the tail and pore region. The states are defined in
	  Table~\ref{table:states}. See Figure~\ref{fig:state_probabilities}
	  for a description of the abbreviation under each state number. The
	  legend in (a) applies to all subplots. (a) Motion is highly
	  anti-correlated in trapped states. As the number of simultaneously
	  influencing trapping mechanisms increases, the Hurst parameter, $H$,
	  decreases. $H$ is highest (closest to Brownian) during transitions 
	  between states (state T). (b) As more trapping mechanisms simultaneously influence solutes,
	  the width of the hop length distribution ($\sigma$) decreases. The
	  largest hops occur when solutes are unbound in the pores. (c) The
	  weight of the hop length distribution's tails, parameterized by
	  $\alpha_h$, increases as more trapping mechanisms influence solutes
	  simultaneously. The transitional hop distributions have among the
	  heaviest tails.}\label{fig:msddm_parameters}
  \end{figure}
